---
title: "MAP_Cleaning_Algorithm_Github"
author: "Jasmine Khan, jasmine.khan@queensu.ca"
date: "25/10/2022"
output: word_document
editor_options: 
  chunk_output_type: console
---

*Table of Contents:*

1) Run Du et al 'clean.iobp' function
1) Run Khan et al 'clean.iobp.icu' function
3) Load data + format
4) Run algorithm + remove gaps
5) Plot MAP traces

```{r}
#set working directory 
setwd()

#libraries
library(data.table)
library(tidyverse)
library(dplyr)
library(padr)
```

#1) Run Du et al 'clean.iobp' function (Du CH, Glick D, Tung A: Error-checking intraoperative arterial line blood pressures. J Clin Monit Comput 2019; 33:407–412). Added 'thresh' argument to function in order to modify parameters.

```{r}
########################################################################################################
#Title: Error-Checking Intraoperative Arterial Line Blood Pressures		
#Journal: Journal of Clinical Monitoring and Computing		
#Authors: Charles H Du, Avery Tung, David Glick		
#Affiliation: University of Chicago Pritzker School of Medicine		
#Contact: cdu@uchicago.edu		
########################################################################################################

#using simulated data, infer the start and end points of the true readings and trim.  Filter out blips.
#takes dataframe bp.trace which has columns minute, sys.trace, dias.trace, MAP.trace
clean.iobp = function(bp.trace, thresh)
{
  #trim leading and tailing NA
  index = which(!is.na(bp.trace$sys.trace) & !is.na(bp.trace$dias.trace) & !is.na(bp.trace$MAP.trace))
  bp.trace = bp.trace[index[1]:index[length(index)],]
  
  if(nrow(bp.trace) < 3)
  {
    print("Aborted: <3 minutes of readings") 
    return(NULL)
  }
  
  #delta is maximum systolic bp change per minute to be considered normal, without artifact or spike
  #try computing a local delta, because at start and end of surgery bp variation is much larger and we 
  #don't want to truncate real peaks
  delta.sys = 8*thresh #original = 8
  delta.dias = 5*thresh #original = 5
  delta.MAP = 6*thresh #original = 6
  SUPPRESS = 3
  
  #infer start as first point which is in physiologic window and holds steady over two minutes
  physiologic = (bp.trace$sys.trace < 260) & (bp.trace$dias.trace < 140) & (bp.trace$sys.trace > 40) & 
                  (bp.trace$dias.trace > 20) & ((bp.trace$sys.trace - bp.trace$dias.trace) > 15)
  
  lshift1s = c(bp.trace$sys.trace[2:nrow(bp.trace)], 0)
  lshift2s = c(bp.trace$sys.trace[3:nrow(bp.trace)], 0, 0)
  lshift1d = c(bp.trace$dias.trace[2:nrow(bp.trace)], 0)
  lshift2d = c(bp.trace$dias.trace[3:nrow(bp.trace)], 0, 0)
  lshift1m = c(bp.trace$MAP.trace[2:nrow(bp.trace)], 0)
  lshift2m = c(bp.trace$MAP.trace[3:nrow(bp.trace)], 0, 0)
  
  #want any two of three (systolic, diastolic, MAP) to hold steady over two minutes, to be more forgiving of errors
  #start and one of next two readings must be physiologic
  index = which(physiologic & (c(physiologic[2:length(physiologic)], T) | c(physiologic[3:length(physiologic)], T, T)) &
                (((abs(lshift1s - bp.trace$sys.trace) <= 2 * delta.sys | abs(lshift2s - bp.trace$sys.trace) <= 2 * sqrt(2)* delta.sys)
                & (abs(lshift1d - bp.trace$dias.trace) <= 2 * delta.dias | abs(lshift2d - bp.trace$dias.trace) <= 2 * sqrt(2)* delta.dias))
                | ((abs(lshift1s - bp.trace$sys.trace) <= 2 * delta.sys | abs(lshift2s - bp.trace$sys.trace) <= 2 * sqrt(2)* delta.sys)
                & (abs(lshift1m - bp.trace$MAP.trace) <= 2 * delta.MAP | abs(lshift2m - bp.trace$MAP.trace) <= 2 * sqrt(2)* delta.MAP))
                | ((abs(lshift1d - bp.trace$dias.trace) <= 2 * delta.dias | abs(lshift2d - bp.trace$dias.trace) <= 2 * sqrt(2)* delta.dias)
                & (abs(lshift1m - bp.trace$MAP.trace) <= 2 * delta.MAP | abs(lshift2m - bp.trace$MAP.trace) <= 2 * sqrt(2)* delta.MAP))))
  inferred.start = index[1]
  
  if(is.na(inferred.start))
  {
    print("Aborted: all readings look artifactual")
    return(NULL)
  }
  
  #infer end as last point which is in physiologic window and holds steady over two minutes
  rshift1s = c(0, bp.trace$sys.trace[1:(nrow(bp.trace) - 1)])
  rshift2s = c(0, 0, bp.trace$sys.trace[1:(nrow(bp.trace) - 2)])
  rshift1d = c(0, bp.trace$dias.trace[1:(nrow(bp.trace) - 1)])
  rshift2d = c(0, 0, bp.trace$dias.trace[1:(nrow(bp.trace) - 2)])
  rshift1m = c(0, bp.trace$MAP.trace[1:(nrow(bp.trace) - 1)])
  rshift2m = c(0, 0, bp.trace$MAP.trace[1:(nrow(bp.trace) - 2)])
  
  index = which(physiologic & (c(T, physiologic[1:(length(physiologic)-1)]) | c(T, T, physiologic[1:(length(physiologic)-2)])) &
                (((abs(rshift1s - bp.trace$sys.trace) <= 2 * delta.sys | abs(rshift2s - bp.trace$sys.trace) <= 2 * sqrt(2)* delta.sys)
                & (abs(rshift1d - bp.trace$dias.trace) <= 2 * delta.dias | abs(rshift2d - bp.trace$dias.trace) <= 2 * sqrt(2)* delta.dias))
                | ((abs(rshift1s - bp.trace$sys.trace) <= 2 * delta.sys | abs(rshift2s - bp.trace$sys.trace) <= 2 * sqrt(2)* delta.sys)
                & (abs(rshift1m - bp.trace$MAP.trace) <= 2 * delta.MAP | abs(rshift2m - bp.trace$MAP.trace) <= 2 * sqrt(2)* delta.MAP))
                | ((abs(rshift1d - bp.trace$dias.trace) <= 2 * delta.dias | abs(rshift2d - bp.trace$dias.trace) <= 2 * sqrt(2)* delta.dias)
                & (abs(rshift1m - bp.trace$MAP.trace) <= 2 * delta.MAP | abs(rshift2m - bp.trace$MAP.trace) <= 2 * sqrt(2)* delta.MAP))))
  inferred.end = index[length(index)]
  
  if(length(inferred.end) == 0)
  {
    print("Aborted: all readings look artifactual2")
    return(NULL)
  }
  
  bp.trace.clean = bp.trace[inferred.start:inferred.end,]
  
  #get average spacing of MAP between systolic and diastolic.  Use this for interpolation of endpoints where linear interpolation is impossible.
  spacings = (bp.trace.clean$MAP.trace - bp.trace.clean$dias.trace) / (bp.trace.clean$sys.trace - bp.trace.clean$dias.trace)
  spacings = spacings[which(spacings > 0 & spacings < 1)]
  spacing = mean(spacings, na.rm = T)
  
  #filter out blips.  Detect by checking if flanking points are both greater or both smaller, and close to one another.
  #keep track of true peaks (in terms of minutes to solve issues with shifting indices).  
  #When a true peak is detected, compute how many fold of delta it is.  Prevent smoothing in other corresponding traces
  #of changes with comparable fold.
  true.peaks.pos = rep(0, nrow(bp.trace.clean) - 1)
  names(true.peaks.pos) = paste(bp.trace.clean$minute[1:(nrow(bp.trace.clean) - 1)], bp.trace.clean$minute[2:nrow(bp.trace.clean)], sep = "-")
  true.peaks.neg = rep(0, nrow(bp.trace.clean) - 1)
  names(true.peaks.neg) = paste(bp.trace.clean$minute[1:(nrow(bp.trace.clean) - 1)], bp.trace.clean$minute[2:nrow(bp.trace.clean)], sep = "-")
    
  #identify systolic points of rapid change.  Track by endpoints. Check if blip or peak.  If blip, Replace with NA.  
  #If peak, store in true.peaks 
  #index = which(abs(diff(bp.trace.clean$sys.trace)) > delta.sys)
  #index = sort(unique(c(index, index+1)))
  #also consider all points bordered by NA, but are not NA
  #missing = which(is.na(bp.trace.clean$sys.trace))
  #to.add = setdiff(c(missing - 1, missing + 1), c(0, nrow(bp.trace.clean)))
  #index = sort(unique(c(index, to.add)))
  #index = setdiff(index, missing)
  index= 1:nrow(bp.trace.clean)
  for(i in index)
  {
    if(is.na(bp.trace.clean$sys.trace[i]))
    {
      next
    }
    
    #if first/last difference is > delta, use 3rd point to check if 1st or 2nd is faulty
    if(i == 1)
    {
      right = 2
      while(is.na(bp.trace.clean$sys.trace[i + right]))
      {
        right = right + 1
      }
      
      if(abs(bp.trace.clean$sys.trace[1 + right] - bp.trace.clean$sys.trace[1]) <= 2 * sqrt(right) * delta.sys)
      {
        next
      }else
      {
        bp.trace.clean$sys.trace[1] = bp.trace.clean$dias.trace[1] + (bp.trace.clean$MAP.trace[1] - bp.trace.clean$dias.trace[1]) / spacing
        next
      }
    }
    if(i == nrow(bp.trace.clean))
    {
      left = 2
      while(is.na(bp.trace.clean$sys.trace[i - left]))
      {
        left = left + 1
      }
      
      if(abs(bp.trace.clean$sys.trace[nrow(bp.trace.clean) - left] - bp.trace.clean$sys.trace[nrow(bp.trace.clean)]) 
         <= 2 * sqrt(left) * delta.sys)
      {
        next
      }else
      {
        bp.trace.clean$sys.trace[nrow(bp.trace.clean)] = bp.trace.clean$dias.trace[nrow(bp.trace.clean)] 
          + (bp.trace.clean$MAP.trace[nrow(bp.trace.clean)] - bp.trace.clean$dias.trace[nrow(bp.trace.clean)]) / spacing
        next
      }
    }
    
    #identify nearest non-NA neighbors
    left = 1
    while(is.na(bp.trace.clean$sys.trace[i - left]))
    {
      left = left + 1
      if(i - left < 1)
      {
        break
      }
    }
    if(i - left < 1)
    {
      next
    }
    
    right = 1
    while(is.na(bp.trace.clean$sys.trace[i + right]))
    {
      right = right + 1
      if(i + right > nrow(bp.trace.clean))
      {
        break
      }
    }
    if(i + right > nrow(bp.trace.clean))
    {
      next
    }
    
    #test if reading is greater or less than both neighbors.  This would indicate blip.
    if(abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) > sqrt(left) * delta.sys &
       abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) > sqrt(right) * delta.sys &
       sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) 
       == sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]))
    {
      bp.trace.clean$sys.trace[i] = NA
      next
    }
    
    #test if start of peak or valley.  If so, flag magnitude.
    if(abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) <= sqrt(left) * delta.sys &
       abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) > sqrt(right) * delta.sys)
    {
      #escape the point to the left of a blip by looking one more point to the right
      if(i + right < nrow(bp.trace.clean))
      {
        right2 = 1
        while(is.na(bp.trace.clean$sys.trace[i + right + right2]))
        {
          right2 = right2 + 1
          if(i + right + right2 > nrow(bp.trace.clean))
          {
            break
          }
        }
        if(i + right + right2 > nrow(bp.trace.clean))
        {
          next
        }
        
        if(abs(bp.trace.clean$sys.trace[i + right] - bp.trace.clean$sys.trace[i + right + right2]) > sqrt(right2) * delta.sys & 
           sign(bp.trace.clean$sys.trace[i + right] - bp.trace.clean$sys.trace[i + right + right2]) 
           == -sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]))
        {
          next
        }
      }
      
      tp.right.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i], bp.trace.clean$minute[i+1], sep = "-"))
      tp.right.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i+right-1], bp.trace.clean$minute[i+right], sep = "-"))
      if(bp.trace.clean$sys.trace[i] < bp.trace.clean$sys.trace[i + right])
      {
        true.peaks.pos[tp.right.start:tp.right.end] = pmax(true.peaks.pos[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) / sqrt(right) / delta.sys, na.rm = T)
        next
      }
      
      if(bp.trace.clean$sys.trace[i] > bp.trace.clean$sys.trace[i + right])
      {
        true.peaks.neg[tp.right.start:tp.right.end] = pmax(true.peaks.neg[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) / sqrt(right) / delta.sys, na.rm = T)
        next
      }
    }
    #test if end of peak or valley.  If so, flag magnitude.
    if(abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) > sqrt(left) * delta.sys &
       abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) <= sqrt(right) * delta.sys)
    {
      #escape the point to the right of a blip by looking one more point to the left
      if(i - left > 1)
      {
        left2 = 1
        while(is.na(bp.trace.clean$sys.trace[i - left - left2]))
        {
          left2 = left2 + 1
          if(i - left - left2 < 1)
          {
            break
          }
        }
        if(i - left - left2 < 1)
        {
          next
        }
        
        if(abs(bp.trace.clean$sys.trace[i - left] - bp.trace.clean$sys.trace[i - left - left2]) > sqrt(left2) * delta.sys & 
           sign(bp.trace.clean$sys.trace[i - left] - bp.trace.clean$sys.trace[i - left - left2]) ==
           -sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]))
        {
          next
        }
      }
      
      tp.left.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-1], bp.trace.clean$minute[i], sep = "-"))
      tp.left.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-left], bp.trace.clean$minute[i-left+1], sep = "-"))
      if(bp.trace.clean$sys.trace[i] > bp.trace.clean$sys.trace[i - left])
      {
        true.peaks.pos[tp.left.start:tp.left.end] = pmax(true.peaks.pos[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) / sqrt(left) / delta.sys, na.rm = T)
        next
      }
      if(bp.trace.clean$sys.trace[i] < bp.trace.clean$sys.trace[i - left])
      {
        true.peaks.neg[tp.left.start:tp.left.end] = pmax(true.peaks.neg[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) / sqrt(left) / delta.sys, na.rm = T)
        next
      }
    }
    
    #test if greater or less than both neighbors but subthreshold, then flag magnitude
    if(sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) 
       == sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]))
    {
      #escape the point to the right of a blip by looking one more point to the left
      if(i - left > 1)
      {
        left2 = 1
        while(is.na(bp.trace.clean$sys.trace[i - left - left2]))
        {
          left2 = left2 + 1
          if(i - left - left2 < 1)
          {
            break
          }
        }
        if(i - left - left2 < 1)
        {
          break
        }
        
        if(abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) > sqrt(left) * delta.sys & 
           abs(bp.trace.clean$sys.trace[i - left] - bp.trace.clean$sys.trace[i - left - left2]) > sqrt(left2) * delta.sys & 
           sign(bp.trace.clean$sys.trace[i - left] - bp.trace.clean$sys.trace[i - left - left2]) ==
           -sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]))
        {
          next
        }
      }
      
      #escape the point to the left of a blip by looking one more point to the right
      if(i + right < nrow(bp.trace.clean))
      {
        right2 = 1
        while(is.na(bp.trace.clean$sys.trace[i + right + right2]))
        {
          right2 = right2 + 1
          if(i + right + right2 > nrow(bp.trace.clean))
          {
            break
          }
        }
        if(i + right + right2 > nrow(bp.trace.clean))
        {
          next
        }
        
        if(abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) > sqrt(right) * delta.sys & 
           abs(bp.trace.clean$sys.trace[i + right] - bp.trace.clean$sys.trace[i + right + right2]) > sqrt(right2) * delta.sys & 
           sign(bp.trace.clean$sys.trace[i + right] - bp.trace.clean$sys.trace[i + right + right2]) 
           == -sign(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]))
        {
          next
        }
      }
      
      tp.right.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i], bp.trace.clean$minute[i+1], sep = "-"))
      tp.right.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i+right-1], bp.trace.clean$minute[i+right], sep = "-"))
      tp.left.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-1], bp.trace.clean$minute[i], sep = "-"))
      tp.left.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-left], bp.trace.clean$minute[i-left+1], sep = "-"))
      
      if(bp.trace.clean$sys.trace[i] > bp.trace.clean$sys.trace[i - left])
      {
        true.peaks.pos[tp.left.start:tp.left.end] = pmax(true.peaks.pos[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) / sqrt(left) / delta.sys, na.rm = T)
        next
      }
      if(bp.trace.clean$sys.trace[i] < bp.trace.clean$sys.trace[i - left])
      {
        true.peaks.neg[tp.left.start:tp.left.end] = pmax(true.peaks.neg[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i - left]) / sqrt(left) / delta.sys, na.rm = T)
        next
      }
      
      if(bp.trace.clean$sys.trace[i] < bp.trace.clean$sys.trace[i + right])
      {
        true.peaks.pos[tp.right.start:tp.right.end] = pmax(true.peaks.pos[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) / sqrt(right) / delta.sys, na.rm = T)
        next
      }
      
      if(bp.trace.clean$sys.trace[i] > bp.trace.clean$sys.trace[i + right])
      {
        true.peaks.neg[tp.right.start:tp.right.end] = pmax(true.peaks.neg[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$sys.trace[i] - bp.trace.clean$sys.trace[i + right]) / sqrt(right) / delta.sys, na.rm = T)
        next
      }
      
    }
  }
  
  #repeat for diastolic
  #index = which(abs(diff(bp.trace.clean$dias.trace)) > delta.dias)
  #index = sort(unique(c(index, index+1)))
  
  #missing = which(is.na(bp.trace.clean$dias.trace))
  #to.add = setdiff(c(missing - 1, missing + 1), c(0, nrow(bp.trace.clean)))
  #index = sort(unique(c(index, to.add)))
  #index = setdiff(index, missing)
  index = 1:nrow(bp.trace.clean)
  for(i in index)
  {
    if(is.na(bp.trace.clean$dias.trace[i]))
    {
      next
    }
    
    #if first/last difference is > delta, use 3rd point to check if 1st or 2nd is faulty
    if(i == 1)
    {
      right = 2
      while(is.na(bp.trace.clean$dias.trace[i + right]))
      {
        right = right + 1
      }
      
      if(abs(bp.trace.clean$dias.trace[1 + right] - bp.trace.clean$dias.trace[1]) <= 2 * sqrt(right) * delta.dias)
      {
        next
      }else
      {
        bp.trace.clean$dias.trace[1] = bp.trace.clean$sys.trace[1] - (bp.trace.clean$sys.trace[1] - bp.trace.clean$MAP.trace[1]) / (1 - spacing)
        next
      }
    }
    if(i == nrow(bp.trace.clean))
    {
      left = 2
      while(is.na(bp.trace.clean$dias.trace[i - left]))
      {
        left = left + 1
      }
      
      if(abs(bp.trace.clean$dias.trace[nrow(bp.trace.clean) - left] - bp.trace.clean$dias.trace[nrow(bp.trace.clean)]) 
         <= 2 * sqrt(left) * delta.dias)
      {
        next
      }else
      {
        bp.trace.clean$dias.trace[nrow(bp.trace.clean)] = bp.trace.clean$sys.trace[nrow(bp.trace.clean)] 
          - (bp.trace.clean$sys.trace[nrow(bp.trace.clean)] - bp.trace.clean$MAP.trace[nrow(bp.trace.clean)]) / (1 - spacing)
        next
      }
    }
    
    #identify nearest non-NA neighbors
    left = 1
    while(is.na(bp.trace.clean$dias.trace[i - left]))
    {
      left = left + 1
      if(i - left < 1)
      {
        break
      }
    }
    if(i - left < 1)
    {
      next
    }
    
    right = 1
    while(is.na(bp.trace.clean$dias.trace[i + right]))
    {
      right = right + 1
      if(i + right > nrow(bp.trace.clean))
      {
        break
      }
    }
    if(i + right > nrow(bp.trace.clean))
    {
      next
    }
    
    #test if reading is greater or less than both neighbors
    if(abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) > sqrt(left) * delta.dias &
       abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) > sqrt(right) * delta.dias &
       sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) 
       == sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]))
    {
      bp.trace.clean$dias.trace[i] = NA
      next
    }
    
    #test if start of peak or valley.  If so, flag magnitude.
    if(abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) <= sqrt(left) * delta.dias &
       abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) > sqrt(right) * delta.dias)
    {
      #escape the point to the left of a blip by looking one more point to the right
      if(i + right < nrow(bp.trace.clean))
      {
        right2 = 1
        while(is.na(bp.trace.clean$dias.trace[i + right + right2]))
        {
          right2 = right2 + 1
          if(i + right + right2 > nrow(bp.trace.clean))
          {
            break
          }
        }
        if(i + right + right2 > nrow(bp.trace.clean))
        {
          next
        }
        
        if(abs(bp.trace.clean$dias.trace[i + right] - bp.trace.clean$dias.trace[i + right + right2]) > sqrt(right2) * delta.dias & 
           sign(bp.trace.clean$dias.trace[i + right] - bp.trace.clean$dias.trace[i + right + right2]) ==
           -sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]))
        {
          next
        }
      }
      
      tp.right.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i], bp.trace.clean$minute[i+1], sep = "-"))
      tp.right.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i+right-1], bp.trace.clean$minute[i+right], sep = "-"))
      
      if(bp.trace.clean$dias.trace[i] < bp.trace.clean$dias.trace[i + right])
      {
        true.peaks.pos[tp.right.start:tp.right.end] = pmax(true.peaks.pos[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) / sqrt(right) / delta.dias, na.rm = T)
      }
      
      if(bp.trace.clean$dias.trace[i] > bp.trace.clean$dias.trace[i + right])
      {
        true.peaks.neg[tp.right.start:tp.right.end] = pmax(true.peaks.neg[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) / sqrt(right) / delta.dias, na.rm = T)
      }
      
    }
    
    #test if end of peak or valley.  If so, flag magnitude.
    if(abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) > sqrt(left) * delta.dias &
       abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) <= sqrt(right) * delta.dias)
    {
      #escape the point to the right of a blip by looking one more point to the left
      if(i - left > 1)
      {
        left2 = 1
        while(is.na(bp.trace.clean$dias.trace[i - left - left2]))
        {
          left2 = left2 + 1
          if(i - left - left2 < 1)
          {
            break
          }
        }
        if(i - left - left2 < 1)
        {
          next
        }
        
        if(abs(bp.trace.clean$dias.trace[i - left] - bp.trace.clean$dias.trace[i - left - left2]) > sqrt(left2) * delta.dias & 
           sign(bp.trace.clean$dias.trace[i - left] - bp.trace.clean$dias.trace[i - left - left2]) ==
           -sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]))
        {
          next
        }
      }
      
      tp.left.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-1], bp.trace.clean$minute[i], sep = "-"))
      tp.left.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-left], bp.trace.clean$minute[i-left+1], sep = "-"))
      
      if(bp.trace.clean$dias.trace[i] > bp.trace.clean$dias.trace[i - left])
      {
        true.peaks.pos[tp.left.start:tp.left.end] = pmax(true.peaks.pos[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) / sqrt(left) / delta.dias, na.rm = T)
      }
      if(bp.trace.clean$dias.trace[i] < bp.trace.clean$dias.trace[i - left])
      {
        true.peaks.neg[tp.left.start:tp.left.end] = pmax(true.peaks.neg[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) / sqrt(left) / delta.dias, na.rm = T)
      }
      
    }
    
    #test if greater or less than both neighbors but subthreshold, then flag magnitude
    if(sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) 
       == sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]))
    {
      #escape the point to the right of a blip by looking one more point to the left
      if(i - left > 1)
      {
        left2 = 1
        while(is.na(bp.trace.clean$dias.trace[i - left - left2]))
        {
          left2 = left2 + 1
          if(i - left - left2 < 1)
          {
            break
          }
        }
        if(i - left - left2 < 1)
        {
          next
        }
        
        if(abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) > sqrt(left) * delta.dias & 
           abs(bp.trace.clean$dias.trace[i - left] - bp.trace.clean$dias.trace[i - left - left2]) > sqrt(left2) * delta.dias & 
           sign(bp.trace.clean$dias.trace[i - left] - bp.trace.clean$dias.trace[i - left - left2]) ==
           -sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]))
        {
          next
        }
      }
      
      #escape the point to the left of a blip by looking one more point to the right
      if(i + right < nrow(bp.trace.clean))
      {
        right2 = 1
        while(is.na(bp.trace.clean$dias.trace[i + right + right2]))
        {
          right2 = right2 + 1
          if(i + right + right2 > nrow(bp.trace.clean))
          {
            break
          }
        }
        if(i + right + right2 > nrow(bp.trace.clean))
        {
          next
        }
        
        if(abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) > sqrt(right) * delta.dias & 
           abs(bp.trace.clean$dias.trace[i + right] - bp.trace.clean$dias.trace[i + right + right2]) > sqrt(right2) * delta.dias & 
           sign(bp.trace.clean$dias.trace[i + right] - bp.trace.clean$dias.trace[i + right + right2]) 
           == -sign(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]))
        {
          next
        }
      }
      
      tp.right.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i], bp.trace.clean$minute[i+1], sep = "-"))
      tp.right.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i+right-1], bp.trace.clean$minute[i+right], sep = "-"))
      tp.left.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-1], bp.trace.clean$minute[i], sep = "-"))
      tp.left.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-left], bp.trace.clean$minute[i-left+1], sep = "-"))
      
      if(bp.trace.clean$dias.trace[i] > bp.trace.clean$dias.trace[i - left])
      {
        true.peaks.pos[tp.left.start:tp.left.end] = pmax(true.peaks.pos[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) / sqrt(left) / delta.dias, na.rm = T)
      }
      if(bp.trace.clean$dias.trace[i] < bp.trace.clean$dias.trace[i - left])
      {
        true.peaks.neg[tp.left.start:tp.left.end] = pmax(true.peaks.neg[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i - left]) / sqrt(left) / delta.dias, na.rm = T)
      }
      
      if(bp.trace.clean$dias.trace[i] < bp.trace.clean$dias.trace[i + right])
      {
        true.peaks.pos[tp.right.start:tp.right.end] = pmax(true.peaks.pos[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) / sqrt(right) / delta.dias, na.rm = T)
      }
      
      if(bp.trace.clean$dias.trace[i] > bp.trace.clean$dias.trace[i + right])
      {
        true.peaks.neg[tp.right.start:tp.right.end] = pmax(true.peaks.neg[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$dias.trace[i] - bp.trace.clean$dias.trace[i + right]) / sqrt(right) / delta.dias, na.rm = T)
      }
      
    }
  }
  
  #repeat for MAP
  #index = which(abs(diff(bp.trace.clean$MAP.trace)) > delta.MAP)
  #index = sort(unique(c(index, index+1)))
  
  #missing = which(is.na(bp.trace.clean$MAP.trace))
  #to.add = setdiff(c(missing - 1, missing + 1), c(0, nrow(bp.trace.clean)))
  #index = sort(unique(c(index, to.add)))
  #index = setdiff(index, missing)
  index = 1:nrow(bp.trace.clean)
  for(i in index)
  {
    if(is.na(bp.trace.clean$MAP.trace[i]))
    {
      next
    }
    
    #if first/last difference is > delta, use 3rd point to check if 1st or 2nd is faulty
    if(i == 1)
    {
      right = 2
      while(is.na(bp.trace.clean$MAP.trace[i + right]))
      {
        right = right + 1
      }
      
      if(abs(bp.trace.clean$MAP.trace[1 + right] - bp.trace.clean$MAP.trace[1]) <= 2 * sqrt(right) * delta.MAP)
      {
        next
      }else
      {
        bp.trace.clean$MAP.trace[1] = bp.trace.clean$dias.trace[1] + (bp.trace.clean$sys.trace[1] - bp.trace.clean$dias.trace[1]) * spacing
        next
      }
    }
    if(i == nrow(bp.trace.clean))
    {
      left = 2
      while(is.na(bp.trace.clean$MAP.trace[i - left]))
      {
        left = left + 1
      }
      
      if(abs(bp.trace.clean$MAP.trace[nrow(bp.trace.clean) - left] - bp.trace.clean$MAP.trace[nrow(bp.trace.clean)]) <= 
         2 * sqrt(left) * delta.MAP)
      {
        next
      }else
      {
        bp.trace.clean$MAP.trace[nrow(bp.trace.clean)] = bp.trace.clean$dias.trace[nrow(bp.trace.clean)] 
          + (bp.trace.clean$sys.trace[nrow(bp.trace.clean)] - bp.trace.clean$dias.trace[nrow(bp.trace.clean)]) * spacing
        next
      }
    }
    
    #identify nearest non-NA neighbors
    left = 1
    while(is.na(bp.trace.clean$MAP.trace[i - left]))
    {
      left = left + 1
      if(i - left < 1)
      {
        break
      }
    }
    if(i - left < 1)
    {
      next
    }
    
    right = 1
    while(is.na(bp.trace.clean$MAP.trace[i + right]))
    {
      right = right + 1
      if(i + right > nrow(bp.trace.clean))
      {
        break
      }
    }
    if(i + right > nrow(bp.trace.clean))
    {
      next
    }
    
    #test if reading is greater or less than both neighbors
    if(abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) > sqrt(left) * delta.MAP &
       abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) > sqrt(right) * delta.MAP &
       sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) == sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]))
    {
      bp.trace.clean$MAP.trace[i] = NA
      next
    }
    
    #test if start of peak or valley.  If so, flag magnitude.
    if(abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) <= sqrt(left) * delta.MAP &
       abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) > sqrt(right) * delta.MAP)
    {
      #escape the point to the left of a blip by looking one more point to the right
      if(i + right < nrow(bp.trace.clean))
      {
        right2 = 1
        while(is.na(bp.trace.clean$MAP.trace[i + right + right2]))
        {
          right2 = right2 + 1
          if(i + right + right2 > nrow(bp.trace.clean))
          {
            break
          }
        }
        if(i + right + right2 > nrow(bp.trace.clean))
        {
          next
        }
        
        if(abs(bp.trace.clean$MAP.trace[i + right] - bp.trace.clean$MAP.trace[i + right + right2]) > sqrt(right2) * delta.MAP & 
           sign(bp.trace.clean$MAP.trace[i + right] - bp.trace.clean$MAP.trace[i + right + right2]) ==
           -sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]))
        {
          next
        }
      }
      
      tp.right.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i], bp.trace.clean$minute[i+1], sep = "-"))
      tp.right.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i+right-1], bp.trace.clean$minute[i+right], sep = "-"))
      
      if(bp.trace.clean$MAP.trace[i] < bp.trace.clean$MAP.trace[i + right])
      {
        true.peaks.pos[tp.right.start:tp.right.end] = pmax(true.peaks.pos[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) / sqrt(right) / delta.MAP, na.rm = T)
      }
      
      if(bp.trace.clean$MAP.trace[i] > bp.trace.clean$MAP.trace[i + right])
      {
        true.peaks.neg[tp.right.start:tp.right.end] = pmax(true.peaks.neg[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) / sqrt(right) / delta.MAP, na.rm = T)
      }
      
    }
    #test if end of peak or valley.  If so, flag magnitude.
    if(abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) > sqrt(left) * delta.MAP &
       abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) <= sqrt(right) * delta.MAP)
    {
      #escape the point to the right of a blip by looking one more point to the left
      if(i - left > 1)
      {
        left2 = 1
        while(is.na(bp.trace.clean$MAP.trace[i - left - left2]))
        {
          left2 = left2 + 1
          if(i - left - left2 < 1)
          {
            break
          }
        }
        if(i - left - left2 < 1)
        {
          next
        }
        
        if(abs(bp.trace.clean$MAP.trace[i - left] - bp.trace.clean$MAP.trace[i - left - left2]) > sqrt(left2) * delta.MAP & 
           sign(bp.trace.clean$MAP.trace[i - left] - bp.trace.clean$MAP.trace[i - left - left2]) ==
           -sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]))
        {
          next
        }
      }
      
      tp.left.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-1], bp.trace.clean$minute[i], sep = "-"))
      tp.left.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-left], bp.trace.clean$minute[i-left+1], sep = "-"))
      
      if(bp.trace.clean$MAP.trace[i] > bp.trace.clean$MAP.trace[i - left])
      {
        true.peaks.pos[tp.left.start:tp.left.end] = pmax(true.peaks.pos[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) / sqrt(left) / delta.MAP, na.rm = T)
      }
      if(bp.trace.clean$MAP.trace[i] < bp.trace.clean$MAP.trace[i - left])
      {
        true.peaks.neg[tp.left.start:tp.left.end] = pmax(true.peaks.neg[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) / sqrt(left) / delta.MAP, na.rm = T)
      }
      
    }
    
    #test if greater or less than both neighbors but subthreshold, then flag magnitude
    if(sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) 
       == sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]))
    {
      #escape the point to the right of a blip by looking one more point to the left
      if(i - left > 1)
      {
        left2 = 1
        while(is.na(bp.trace.clean$MAP.trace[i - left - left2]))
        {
          left2 = left2 + 1
          if(i - left - left2 < 1)
          {
            break
          }
        }
        if(i - left - left2 < 1)
        {
          next
        }
        
        if(abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) > sqrt(left) * delta.MAP & 
           abs(bp.trace.clean$MAP.trace[i - left] - bp.trace.clean$MAP.trace[i - left - left2]) > sqrt(left2) * delta.MAP & 
           sign(bp.trace.clean$MAP.trace[i - left] - bp.trace.clean$MAP.trace[i - left - left2]) ==
           -sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]))
        {
          next
        }
      }
      
      #escape the point to the left of a blip by looking one more point to the right
      if(i + right < nrow(bp.trace.clean))
      {
        right2 = 1
        while(is.na(bp.trace.clean$MAP.trace[i + right + right2]))
        {
          right2 = right2 + 1
          if(i + right + right2 > nrow(bp.trace.clean))
          {
            break
          }
        }
        if(i + right + right2 > nrow(bp.trace.clean))
        {
          next
        }
        
        if(abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) > sqrt(right) * delta.MAP & 
           abs(bp.trace.clean$MAP.trace[i + right] - bp.trace.clean$MAP.trace[i + right + right2]) > sqrt(right2) * delta.MAP & 
           sign(bp.trace.clean$MAP.trace[i + right] - bp.trace.clean$MAP.trace[i + right + right2]) 
           == -sign(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]))
        {
          next
        }
      }
      
      tp.right.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i], bp.trace.clean$minute[i+1], sep = "-"))
      tp.right.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i+right-1], bp.trace.clean$minute[i+right], sep = "-"))
      tp.left.start = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-1], bp.trace.clean$minute[i], sep = "-"))
      tp.left.end = which(names(true.peaks.pos) == paste(bp.trace.clean$minute[i-left], bp.trace.clean$minute[i-left+1], sep = "-"))
      
      if(bp.trace.clean$MAP.trace[i] > bp.trace.clean$MAP.trace[i - left])
      {
        true.peaks.pos[tp.left.start:tp.left.end] = pmax(true.peaks.pos[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) / sqrt(left) / delta.MAP, na.rm = T)
      }
      if(bp.trace.clean$MAP.trace[i] < bp.trace.clean$MAP.trace[i - left])
      {
        true.peaks.neg[tp.left.start:tp.left.end] = pmax(true.peaks.neg[tp.left.start:tp.left.end],
                                                         abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i - left]) / sqrt(left) / delta.MAP, na.rm = T)
      }
      
      if(bp.trace.clean$MAP.trace[i] < bp.trace.clean$MAP.trace[i + right])
      {
        true.peaks.pos[tp.right.start:tp.right.end] = pmax(true.peaks.pos[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) / sqrt(right) / delta.MAP, na.rm = T)
      }
      
      if(bp.trace.clean$MAP.trace[i] > bp.trace.clean$MAP.trace[i + right])
      {
        true.peaks.neg[tp.right.start:tp.right.end] = pmax(true.peaks.neg[tp.right.start:tp.right.end],
                                                           abs(bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i + right]) / sqrt(right) / delta.MAP, na.rm = T)
      }
      
      
    }
  }
  
  
  #fill back in original values for each trace if the fold change relative to delta is less than SUPPRESS * true.peaks
  index = which(is.na(bp.trace.clean$sys.trace))
  for(i in index)
  {
    #find j, corresponding index in bp.trace
    time = bp.trace.clean$minute[i]
    j = which(bp.trace$minute == time)
    
    #nothing to do if originally NA
    if(is.na(bp.trace$sys.trace[j]))
    {
      next
    }
    
    #handle possible flanking NAs
    left = 1
    while(is.na(bp.trace$sys.trace[j - left]))
    {
      left = left + 1
    }
    
    right = 1
    while(is.na(bp.trace$sys.trace[j + right]))
    {
      right = right + 1
    }
    
    #check rates of change on both sides, add originally value back if suppressed by a larger true peak in another trace.
    if(bp.trace$sys.trace[j] > bp.trace$sys.trace[j - left])
    {
      if(abs(bp.trace$sys.trace[j] - bp.trace$sys.trace[j - left]) / delta.sys / sqrt(left) < SUPPRESS * true.peaks.pos[paste(time - 1, time, sep = "-")])
      {
        bp.trace.clean$sys.trace[i] = bp.trace$sys.trace[j]
      }
    }
    if(bp.trace$sys.trace[j] < bp.trace$sys.trace[j - left])
    {
      if(abs(bp.trace$sys.trace[j] - bp.trace$sys.trace[j - left]) / delta.sys / sqrt(left) < SUPPRESS * true.peaks.neg[paste(time - 1, time, sep = "-")])
      {
        bp.trace.clean$sys.trace[i] = bp.trace$sys.trace[j]
      }
    }
    if(bp.trace$sys.trace[j] < bp.trace$sys.trace[j + right])
    {
      if(abs(bp.trace$sys.trace[j] - bp.trace$sys.trace[j + right]) / delta.sys / sqrt(right) < SUPPRESS * true.peaks.pos[paste(time, time + 1, sep = "-")])
      {
        bp.trace.clean$sys.trace[i] = bp.trace$sys.trace[j]
      }
    }
    if(bp.trace$sys.trace[j] > bp.trace$sys.trace[j + right])
    {
      if(abs(bp.trace$sys.trace[j] - bp.trace$sys.trace[j + right]) / delta.sys / sqrt(right) < SUPPRESS * true.peaks.neg[paste(time, time + 1, sep = "-")])
      {
        bp.trace.clean$sys.trace[i] = bp.trace$sys.trace[j]
      }
    }
    
  }
  
  #repeat for diastolic
  index = which(is.na(bp.trace.clean$dias.trace))
  for(i in index)
  {
    #find j, corresponding index in bp.trace
    time = bp.trace.clean$minute[i]
    j = which(bp.trace$minute == time)
    
    #nothing to do if originally NA
    if(is.na(bp.trace$dias.trace[j]))
    {
      next
    }
    
    #handle possible flanking NAs
    left = 1
    while(is.na(bp.trace$dias.trace[j - left]))
    {
      left = left + 1
    }
    
    right = 1
    while(is.na(bp.trace$dias.trace[j + right]))
    {
      right = right + 1
    }
    
    #check rates of change on both sides, add originally value back if suppressed by a larger true peak in another trace.
    if(bp.trace$dias.trace[j] > bp.trace$dias.trace[j - left])
    {
      if(abs(bp.trace$dias.trace[j] - bp.trace$dias.trace[j - left]) / delta.dias / sqrt(left) < SUPPRESS * true.peaks.pos[paste(time - 1, time, sep = "-")])
      {
        bp.trace.clean$dias.trace[i] = bp.trace$dias.trace[j]
      }
    }
    if(bp.trace$dias.trace[j] < bp.trace$dias.trace[j - left])
    {
      if(abs(bp.trace$dias.trace[j] - bp.trace$dias.trace[j - left]) / delta.dias / sqrt(left) < SUPPRESS * true.peaks.neg[paste(time - 1, time, sep = "-")])
      {
        bp.trace.clean$dias.trace[i] = bp.trace$dias.trace[j]
      }
    }
    if(bp.trace$dias.trace[j] < bp.trace$dias.trace[j + right])
    {
      if(abs(bp.trace$dias.trace[j] - bp.trace$dias.trace[j + right]) / delta.dias / sqrt(right) < SUPPRESS * true.peaks.pos[paste(time, time + 1, sep = "-")])
      {
        bp.trace.clean$dias.trace[i] = bp.trace$dias.trace[j]
      }
    }
    if(bp.trace$dias.trace[j] > bp.trace$dias.trace[j + right])
    {
      if(abs(bp.trace$dias.trace[j] - bp.trace$dias.trace[j + right]) / delta.dias / sqrt(right) < SUPPRESS * true.peaks.neg[paste(time, time + 1, sep = "-")])
      {
        bp.trace.clean$dias.trace[i] = bp.trace$dias.trace[j]
      }
    }
    
  }
  
  #repeat for MAP
  index = which(is.na(bp.trace.clean$MAP.trace))
  for(i in index)
  {
    #find j, corresponding index in bp.trace
    time = bp.trace.clean$minute[i]
    j = which(bp.trace$minute == time)
    
    #nothing to do if originally NA
    if(is.na(bp.trace$MAP.trace[j]))
    {
      next
    }
    
    #handle possible flanking NAs
    left = 1
    while(is.na(bp.trace$MAP.trace[j - left]))
    {
      left = left + 1
    }
    
    right = 1
    while(is.na(bp.trace$MAP.trace[j + right]))
    {
      right = right + 1
    }
    
    #check rates of change on both sides, add originally value back if suppressed by a larger true peak in another trace.
    if(bp.trace$MAP.trace[j] > bp.trace$MAP.trace[j - left])
    {
      if(abs(bp.trace$MAP.trace[j] - bp.trace$MAP.trace[j - left]) / delta.MAP / sqrt(left) < SUPPRESS * true.peaks.pos[paste(time - 1, time, sep = "-")])
      {
        bp.trace.clean$MAP.trace[i] = bp.trace$MAP.trace[j]
      }
    }
    if(bp.trace$MAP.trace[j] < bp.trace$MAP.trace[j - left])
    {
      if(abs(bp.trace$MAP.trace[j] - bp.trace$MAP.trace[j - left]) / delta.MAP / sqrt(left) < SUPPRESS * true.peaks.neg[paste(time - 1, time, sep = "-")])
      {
        bp.trace.clean$MAP.trace[i] = bp.trace$MAP.trace[j]
      }
    }
    if(bp.trace$MAP.trace[j] < bp.trace$MAP.trace[j + right])
    {
      if(abs(bp.trace$MAP.trace[j] - bp.trace$MAP.trace[j + right]) / delta.MAP / sqrt(right) < SUPPRESS * true.peaks.pos[paste(time, time + 1, sep = "-")])
      {
        bp.trace.clean$MAP.trace[i] = bp.trace$MAP.trace[j]
      }
    }
    if(bp.trace$MAP.trace[j] > bp.trace$MAP.trace[j + right])
    {
      if(abs(bp.trace$MAP.trace[j] - bp.trace$MAP.trace[j + right]) / delta.MAP / sqrt(right) < SUPPRESS * true.peaks.neg[paste(time, time + 1, sep = "-")])
      {
        bp.trace.clean$MAP.trace[i] = bp.trace$MAP.trace[j]
      }
    }
    
  }
  
  
  #fill in missing values with linear interpolation
  index = which(is.na(bp.trace.clean$sys.trace))
  for(i in index)
  {
    j = i+1
    count = 1
    while(j %in% index)
    {
      index = setdiff(index, j)
      count = count + 1
      j = j + 1
    }
    bp.trace.clean$sys.trace[i:(i + count - 1)] = rep(bp.trace.clean$sys.trace[i-1], count) + 1:count * 
      (bp.trace.clean$sys.trace[i + count] - bp.trace.clean$sys.trace[i - 1]) / (count + 1)
  }
  #repeat for diastolic
  index = which(is.na(bp.trace.clean$dias.trace))
  for(i in index)
  {
    j = i+1
    count = 1
    while(j %in% index)
    {
      index = setdiff(index, j)
      count = count + 1
      j = j + 1
    }
    bp.trace.clean$dias.trace[i:(i + count - 1)] = rep(bp.trace.clean$dias.trace[i-1], count) + 1:count * 
      (bp.trace.clean$dias.trace[i + count] - bp.trace.clean$dias.trace[i - 1]) / (count + 1)
  }
  #repeat for MAP
  index = which(is.na(bp.trace.clean$MAP.trace))
  for(i in index)
  {
    j = i+1
    count = 1
    while(j %in% index)
    {
      index = setdiff(index, j)
      count = count + 1
      j = j + 1
    }
    bp.trace.clean$MAP.trace[i:(i + count - 1)] = rep(bp.trace.clean$MAP.trace[i-1], count) + 1:count * 
      (bp.trace.clean$MAP.trace[i + count] - bp.trace.clean$MAP.trace[i - 1]) / (count + 1)
  }
  
  #cross reference systolic, diastolic, and MAP for consistency
  #first fetch instances where MAP is too high, or sys or dias is too low
  index = which(bp.trace.clean$MAP.trace > bp.trace.clean$dias.trace + (1 - 2/3 * (1-spacing)) * (bp.trace.clean$sys.trace - bp.trace.clean$dias.trace))
  z = nrow(bp.trace.clean)
  for(i in index)
  {
    #if at endpoints, identify which two of systolic, diastolic, MAP are correct and use them to interpolate the third
    if(i == 1)
    {
      if(max(c(bp.trace.clean$sys.trace[2] - bp.trace.clean$sys.trace[1], bp.trace.clean$dias.trace[2] - bp.trace.clean$dias.trace[1], 
               bp.trace.clean$MAP.trace[1] - bp.trace.clean$MAP.trace[2])) == bp.trace.clean$sys.trace[2] - bp.trace.clean$sys.trace[1])
      {
        bp.trace.clean$sys.trace[1] = bp.trace.clean$dias.trace[1] + (bp.trace.clean$MAP.trace[1] - bp.trace.clean$dias.trace[1]) / spacing
      }
      if(max(c(bp.trace.clean$sys.trace[2] - bp.trace.clean$sys.trace[1], bp.trace.clean$dias.trace[2] - bp.trace.clean$dias.trace[1], 
               bp.trace.clean$MAP.trace[1] - bp.trace.clean$MAP.trace[2])) == bp.trace.clean$dias.trace[2] - bp.trace.clean$dias.trace[1])
      {
        bp.trace.clean$dias.trace[1] = bp.trace.clean$sys.trace[1] - (bp.trace.clean$sys.trace[1] - bp.trace.clean$MAP.trace[1]) / (1 - spacing)
      }
      if(max(c(bp.trace.clean$sys.trace[2] - bp.trace.clean$sys.trace[1], bp.trace.clean$dias.trace[2] - bp.trace.clean$dias.trace[1], 
               bp.trace.clean$MAP.trace[1] - bp.trace.clean$MAP.trace[2])) == bp.trace.clean$MAP.trace[1] - bp.trace.clean$MAP.trace[2])
      {
        bp.trace.clean$MAP.trace[1] = bp.trace.clean$dias.trace[1] + (bp.trace.clean$sys.trace[1] - bp.trace.clean$dias.trace[1]) * spacing
      }
      next
    }
    if(i == z)
    {
      if(max(c(bp.trace.clean$sys.trace[z-1] - bp.trace.clean$sys.trace[z], bp.trace.clean$dias.trace[z-1] - bp.trace.clean$dias.trace[z], 
               bp.trace.clean$MAP.trace[z] - bp.trace.clean$MAP.trace[z-1])) == bp.trace.clean$sys.trace[z-1] - bp.trace.clean$sys.trace[z])
      {
        bp.trace.clean$sys.trace[z] = bp.trace.clean$dias.trace[z] + (bp.trace.clean$MAP.trace[z] - bp.trace.clean$dias.trace[z]) / spacing
      }
      if(max(c(bp.trace.clean$sys.trace[z-1] - bp.trace.clean$sys.trace[z], bp.trace.clean$dias.trace[z-1] - bp.trace.clean$dias.trace[z], 
               bp.trace.clean$MAP.trace[z] - bp.trace.clean$MAP.trace[z-1])) == bp.trace.clean$dias.trace[z-1] - bp.trace.clean$dias.trace[z])
      {
        bp.trace.clean$dias.trace[z] = bp.trace.clean$sys.trace[z] - (bp.trace.clean$sys.trace[z] - bp.trace.clean$MAP.trace[z]) / (1 - spacing)
      }
      if(max(c(bp.trace.clean$sys.trace[z-1] - bp.trace.clean$sys.trace[z], bp.trace.clean$dias.trace[z-1] - bp.trace.clean$dias.trace[z], 
               bp.trace.clean$MAP.trace[z] - bp.trace.clean$MAP.trace[z-1])) == bp.trace.clean$MAP.trace[z] - bp.trace.clean$MAP.trace[z-1])
      {
        bp.trace.clean$MAP.trace[z] = bp.trace.clean$dias.trace[z] + (bp.trace.clean$sys.trace[z] - bp.trace.clean$dias.trace[z]) * spacing
      }
      next
    }else #if in the middle of trace, identify aberrant trace and use linear interpolation
    {
      if(max(c(bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i],
               bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i],
               bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1]))
               == bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i])
      {
        #bp.trace.clean$sys.trace[i] = (bp.trace.clean$sys.trace[i-1] + bp.trace.clean$sys.trace[i+1]) / 2
        bp.trace.clean$sys.trace[i] = bp.trace.clean$dias.trace[i] + (bp.trace.clean$MAP.trace[i] - bp.trace.clean$dias.trace[i]) / spacing
      }
      if(max(c(bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i],
               bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i],
               bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1]))
         == bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i])
      {
        #bp.trace.clean$dias.trace[i] = (bp.trace.clean$dias.trace[i-1] + bp.trace.clean$dias.trace[i+1]) / 2
        bp.trace.clean$dias.trace[i] = bp.trace.clean$sys.trace[i] - (bp.trace.clean$sys.trace[i] - bp.trace.clean$MAP.trace[i]) / (1 - spacing)
      }
      if(max(c(bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i],
               bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i],
               bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1]))
         == bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1])
      {
        #bp.trace.clean$MAP.trace[i] = (bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i+1]) / 2
        bp.trace.clean$MAP.trace[i] = bp.trace.clean$dias.trace[i] + (bp.trace.clean$sys.trace[i] - bp.trace.clean$dias.trace[i]) * spacing
      }
    }
  }
  #next fetch instances where MAP.trace is too low, or sys or dias too high
  index = which(bp.trace.clean$MAP.trace < bp.trace.clean$dias.trace + 2/3 * spacing * (bp.trace.clean$sys.trace - bp.trace.clean$dias.trace))
  for(i in index)
  {
    #if at endpoints, identify which two of systolic, diastolic, MAP are correct and use them to interpolate the third
    if(i == 1)
    {
      if(max(c(bp.trace.clean$sys.trace[1] - bp.trace.clean$sys.trace[2], bp.trace.clean$dias.trace[1] - bp.trace.clean$dias.trace[2], 
               bp.trace.clean$MAP.trace[2] - bp.trace.clean$MAP.trace[1])) == bp.trace.clean$sys.trace[1] - bp.trace.clean$sys.trace[2])
      {
        bp.trace.clean$sys.trace[1] = bp.trace.clean$dias.trace[1] + (bp.trace.clean$MAP.trace[1] - bp.trace.clean$dias.trace[1]) / spacing
      }
      if(max(c(bp.trace.clean$sys.trace[1] - bp.trace.clean$sys.trace[2], bp.trace.clean$dias.trace[1] - bp.trace.clean$dias.trace[2], 
               bp.trace.clean$MAP.trace[2] - bp.trace.clean$MAP.trace[1])) == bp.trace.clean$dias.trace[1] - bp.trace.clean$dias.trace[2])
      {
        bp.trace.clean$dias.trace[1] = bp.trace.clean$sys.trace[1] - (bp.trace.clean$sys.trace[1] - bp.trace.clean$MAP.trace[1]) / (1 - spacing)
      }
      if(max(c(bp.trace.clean$sys.trace[1] - bp.trace.clean$sys.trace[2], bp.trace.clean$dias.trace[1] - bp.trace.clean$dias.trace[2], 
               bp.trace.clean$MAP.trace[2] - bp.trace.clean$MAP.trace[1])) == bp.trace.clean$MAP.trace[2] - bp.trace.clean$MAP.trace[1])
      {
        bp.trace.clean$MAP.trace[1] = bp.trace.clean$dias.trace[1] + (bp.trace.clean$sys.trace[1] - bp.trace.clean$dias.trace[1]) * spacing
      }
      next
    }
    if(i == z)
    {
      if(max(c(bp.trace.clean$sys.trace[z] - bp.trace.clean$sys.trace[z-1], bp.trace.clean$dias.trace[z] - bp.trace.clean$dias.trace[z-1], 
               bp.trace.clean$MAP.trace[z-1] - bp.trace.clean$MAP.trace[z])) == bp.trace.clean$sys.trace[z] - bp.trace.clean$sys.trace[z-1])
      {
        bp.trace.clean$sys.trace[z] = bp.trace.clean$dias.trace[z] + (bp.trace.clean$MAP.trace[z] - bp.trace.clean$dias.trace[z]) / spacing
      }
      if(max(c(bp.trace.clean$sys.trace[z] - bp.trace.clean$sys.trace[z-1], bp.trace.clean$dias.trace[z] - bp.trace.clean$dias.trace[z-1], 
               bp.trace.clean$MAP.trace[z-1] - bp.trace.clean$MAP.trace[z])) == bp.trace.clean$dias.trace[z] - bp.trace.clean$dias.trace[z-1])
      {
        bp.trace.clean$dias.trace[z] = bp.trace.clean$sys.trace[z] - (bp.trace.clean$sys.trace[z] - bp.trace.clean$MAP.trace[z]) / (1 - spacing)
      }
      if(max(c(bp.trace.clean$sys.trace[z] - bp.trace.clean$sys.trace[z-1], bp.trace.clean$dias.trace[z] - bp.trace.clean$dias.trace[z-1], 
               bp.trace.clean$MAP.trace[z-1] - bp.trace.clean$MAP.trace[z])) == bp.trace.clean$MAP.trace[z-1] - bp.trace.clean$MAP.trace[z])
      {
        bp.trace.clean$MAP.trace[z] = bp.trace.clean$dias.trace[z] + (bp.trace.clean$sys.trace[z] - bp.trace.clean$dias.trace[z]) * spacing
      }
      next
    }else #if in the middle of trace, identify aberrant trace and use linear interpolation
    {
      if(min(c(bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i],
               bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i],
               bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1]))
         == bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i])
      {
        #bp.trace.clean$sys.trace[i] = (bp.trace.clean$sys.trace[i-1] + bp.trace.clean$sys.trace[i+1]) / 2
        bp.trace.clean$sys.trace[i] = bp.trace.clean$dias.trace[i] + (bp.trace.clean$MAP.trace[i] - bp.trace.clean$dias.trace[i]) / spacing
      }
      if(min(c(bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i],
               bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i],
               bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1]))
         == bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i])
      {
        #bp.trace.clean$dias.trace[i] = (bp.trace.clean$dias.trace[i-1] + bp.trace.clean$dias.trace[i+1]) / 2
        bp.trace.clean$dias.trace[i] = bp.trace.clean$sys.trace[i] - (bp.trace.clean$sys.trace[i] - bp.trace.clean$MAP.trace[i]) / (1 - spacing)
      }
      if(min(c(bp.trace.clean$sys.trace[i-1] - bp.trace.clean$sys.trace[i] + bp.trace.clean$sys.trace[i+1] - bp.trace.clean$sys.trace[i],
               bp.trace.clean$dias.trace[i-1] - bp.trace.clean$dias.trace[i] + bp.trace.clean$dias.trace[i+1] - bp.trace.clean$dias.trace[i],
               bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1]))
         == bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i] - bp.trace.clean$MAP.trace[i+1])
      {
        #bp.trace.clean$MAP.trace[i] = (bp.trace.clean$MAP.trace[i-1] + bp.trace.clean$MAP.trace[i+1]) / 2
        bp.trace.clean$MAP.trace[i] = bp.trace.clean$dias.trace[i] + (bp.trace.clean$sys.trace[i] - bp.trace.clean$dias.trace[i]) * spacing
      }
    }
  }
  
  #finally check for improperly tight spacings and interpolate linearly
  index.sd = (bp.trace.clean$sys.trace - bp.trace.clean$dias.trace) < pmin(bp.trace.clean$sys.trace/5, 15)
  index.sm = (bp.trace.clean$sys.trace - bp.trace.clean$MAP.trace) < pmin(bp.trace.clean$sys.trace/9, 8)
  index.md = (bp.trace.clean$MAP.trace - bp.trace.clean$dias.trace) < pmin(bp.trace.clean$MAP.trace/12, 5)
  
  bp.trace.clean$sys.trace[index.sd | index.sm] = NA
  bp.trace.clean$dias.trace[index.sd | index.md] = NA
  bp.trace.clean$MAP.trace[index.sm | index.md] = NA
  
  #fill in missing values from final spacing check with linear interpolation
  #systolic
  index = which(is.na(bp.trace.clean$sys.trace))
  for(i in index)
  {
    if((i == 1)|(i == nrow(bp.trace.clean)))
    {
      bp.trace.clean = bp.trace.clean[setdiff(1:nrow(bp.trace.clean), i),] 
      next
    }
    
    j = i+1
    count = 1
    while(j %in% index)
    {
      index = setdiff(index, j)
      count = count + 1
      j = j + 1
    }
    bp.trace.clean$sys.trace[i:(i + count - 1)] = rep(bp.trace.clean$sys.trace[i-1], count) + 1:count * 
      (bp.trace.clean$sys.trace[i + count] - bp.trace.clean$sys.trace[i - 1]) / (count + 1)
  }
  #repeat for diastolic
  index = which(is.na(bp.trace.clean$dias.trace))
  for(i in index)
  {
    if((i == 1)|(i == nrow(bp.trace.clean)))
    {
      bp.trace.clean = bp.trace.clean[setdiff(1:nrow(bp.trace.clean), i),] 
      next
    }
    
    j = i+1
    count = 1
    while(j %in% index)
    {
      index = setdiff(index, j)
      count = count + 1
      j = j + 1
    }
    bp.trace.clean$dias.trace[i:(i + count - 1)] = rep(bp.trace.clean$dias.trace[i-1], count) + 1:count * 
      (bp.trace.clean$dias.trace[i + count] - bp.trace.clean$dias.trace[i - 1]) / (count + 1)
  }
  #repeat for MAP
  index = which(is.na(bp.trace.clean$MAP.trace))
  for(i in index)
  {
    if((i == 1)|(i == nrow(bp.trace.clean)))
    {
      bp.trace.clean = bp.trace.clean[setdiff(1:nrow(bp.trace.clean), i),] 
      next
    }
    
    j = i+1
    count = 1
    while(j %in% index)
    {
      index = setdiff(index, j)
      count = count + 1
      j = j + 1
    }
    bp.trace.clean$MAP.trace[i:(i + count - 1)] = rep(bp.trace.clean$MAP.trace[i-1], count) + 1:count * 
      (bp.trace.clean$MAP.trace[i + count] - bp.trace.clean$MAP.trace[i - 1]) / (count + 1)
  }
  
  
  return(bp.trace.clean)
}


```

#2) Run Khan et al 'clean.iobp.icu' function using Du et al 'clean.iobp' function
```{r}
#include in data: subject_id, DateTime, MAP.trace, sys.trace, dias.trace, minute

clean.iobp.icu <- function(bp.trace, lower_limit, upper_limit, threshold) {

  subject.list <- unique(bp.trace$subject_id)
  
  #run through given threshold(s) (1.3 recommended in ICU)
  for (j in 1:length(threshold)) {
    
    print(threshold[j])
    pt.clean <- list()
    pt.raw <- list()
    pt.mins <- NULL
    
    #run Du et al., algorithm on each subject
    for (i in subject.list) {
    
      pt.mins[i] <- as.numeric(length(bp.trace$subject_id[bp.trace$subject_id == i]))
      pt.raw[[i]] = bp.trace[bp.trace$subject_id == i,] 
      pt.clean[[i]] = clean.iobp(pt.raw[[i]],threshold[j])
      
      print(i)
      
    }
  
    #combine patient data into one data frame
    pt.clean <- lapply(pt.clean, mutate, cleaned_minute = row_number()-1)
    algorithm.pt.clean = plyr::ldply(pt.clean)
    
    #clean up data to prep for post-hoc edits
    bp.algorithm <- algorithm.pt.clean %>% 
      select(c("subject_id", "DateTime", "MAP.trace", "sys.trace","dias.trace", "minute", "cleaned_minute")) %>%
      rename("MAP.new" = "MAP.trace", 
             "sys.new" = "sys.trace",
             "dias.new" = "dias.trace")
    
    #post-hoc algorithm edits, removing any values outside physiologic range (<30 and >200mmHg limits recommended for ICU)
    bp.algorithm <- bp.algorithm %>%
      filter(MAP.new > lower_limit,
             MAP.new < upper_limit)
    
  }

  return(bp.algorithm)
}
```

#3.a) Load data and format:
```{r}
bp.original <- fread("base_unpadded.csv", na.strings= c("", "NaN")) 
bp.original <- bp.original %>% 
  select(c("subject_id", 
           "DateTime", 
           "MAP.trace", 
           "sys.trace", 
           "dias.trace"))

bp.no.gaps <- bp.original %>% 
  filter(!(is.na(MAP.trace) & is.na(sys.trace) & is.na(dias.trace))) %>%
  select(subject_id, DateTime)
```

#b) Add minute column and insert any missing timestamps
```{r}
bp.original.pad <- bp.original %>%
  group_by(subject_id) %>%
  pad() %>%
  dplyr::mutate(minute = row_number()-1) 
```

#4) Run algorithm
```{r}
#run algorithm
bp.algorithm <-  clean.iobp.icu(bp.original.pad, 30, 200, 1.3)

#Remove data breaks to remove inappropriately interpolated data points
bp.algorithm.final <- inner_join(bp.algorithm, bp.no.gaps, by = c("subject_id", "DateTime"))
```

#5) Plot original and clean bp traces
```{r}
subject.list <- unique(bp.original$subject_id)

for (i in subject.list) {
  
  og.temp <- filter(bp.original.pad, subject_id == i) 
  alg.temp <- filter(bp.algorithm.final, subject_id == i) %>% pad() #pad so that ggplot doesn't join data breaks
  
  print(ggplot(og.temp, aes(x = minute, y= MAP.trace)) +
  geom_line(color = "red") +
  geom_path(data = alg.temp, color = "#0072B2", na.rm = FALSE, aes(x= minute, y = MAP.new)) + 
  labs(title = i, x= "Time (minutes)", y="MAP (mmHg)") +
  theme_classic() 
  )
  
  remove(og.temp, alg.temp)
}
```

